### 2017-11-30 doc/doc.tex
documentation

\documentclass[a4paper,german]{report}

    \title{Projekt Medieninformatik}
    \date{WS17/18}
    \author{Lewe Ohlsen\\ Fachhochschule Wedel}

    \usepackage[utf8]{inputenc}                   % latin9 utf8
    \usepackage[german]{babel}
    \usepackage{tikz}
    \usepackage{pgfgantt}

    \begin{document}
        \maketitle
            \section{Section Headings}
            \begin{ganttchart}{1}{16}
                \gantttitle{Kalenderwoche}{16} \\
                \gantttitlelist{46,...,52,1,2,3,4,5,6,7,8,9}{1} \\

                \ganttbar{MNIST Tutorial}{1}{2} \\
                \ganttlinkedbar{Datenaufbereitung}{3}{3} \\
                \ganttbar{CNN-Inplementierung}{3}{9} \\

                %\ganttlinkedbar{Task 2}{3}{7} \ganttnewline
                \ganttmilestone{Webcam-Input}{7} \ganttnewline
                \ganttbar{Final Task}{8}{12}

                %\ganttlink{elem1}{elem2}

            \end{ganttchart}
    \end{document}