### 2017-12-11 doc/doc.tex
notes

\documentclass[a4paper,english]{report}

    \title{Projekt Medieninformatik\\Facial Expresson Recognition Using A State Of The Art CNN}
    \date{WS17/18}
    \author{Lewe Ohlsen\\ Fachhochschule Wedel}

    \usepackage[utf8]{inputenc}                   % latin9 utf8
    \usepackage{tikz}
    \usepackage{pgfgantt}

    \begin{document}
        \maketitle
        \newpage
            \tableofcontents
        \newpage
        \section{Introduction}
            In order to get started with deep learning techniques and state of the art neural networks, I am participating in a 2013 kaggle competition on image classification.
        \section{Roadmap}
            \begin{ganttchart}{1}{16}
                \gantttitle{Kalenderwoche}{16} \\
                \gantttitlelist{46,...,52,1,2,3,4,5,6,7,8,9}{1} \\

                \ganttbar{MNIST Tutorial}{1}{2} \\
                \ganttlinkedbar{Data preparation}{3}{3} \\
                \ganttbar{implementing a simple linear model}{3}{4} \\                                
                \ganttlinkedbar{implementing a CNN}{4}{10} \\

                %\ganttlinkedbar{Task 2}{3}{7} \ganttnewline
                \ganttmilestone{processing Webcam images}{8} \ganttnewline
                \ganttlinkedbar{Tuning parameters}{10}{12}

                %\ganttlink{elem1}{elem2}

            \end{ganttchart}
    \end{document}