### 2018-04-15 doc/paper.tex
corrections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Facial Expresion Recognition with Convolutional Neural Networks
% 
% Lewe Ohlsen
% Fachhochschule Wedel
%
% Latex Template by Frits Wenneker (http://www.howtotex.com)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

\usepackage{lipsum} % Package to generate dummy text throughout this template
\usepackage{graphicx} % Figures from matplotlib
\graphicspath{ {figures/} }

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures

\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage{cite} % BibTex

\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\usepackage[utf8]{inputenc}
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering\bfseries}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{}{\thesubsection.}{1em}{} % Change the look of the section titles

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{16pt}{10pt}\selectfont\textbf{Facial Expression Recognition with\\ Convolutional Neural Networks}} % Article title

\author{
	\large
	\textsc{Lewe Ohlsen}\\[2mm] % Your name
	\normalsize Fachhochschule Wedel \\ % Your institution
	\normalsize \href{mailto:minf101062@fh-wedel.de}{minf101062@fh-wedel.de} % Your email address
	\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

\noindent El barómetro es un instrumento de medición atmosérica, específicamente utilizado en la determinación de la fuerza por unidad de superficie ejercida por el peso de la atmósfera. Existe un gran número de equipos atmoféricos con distintos tipos de estos aparátos y son diariamente utilizados ya que la presión atmosférica juega un papel importante en la determinación y pronóstico del tiempo así como en el área de investigación al momento de realizarse experimentos ya que pueden llegar a afectar o hacer variar el funcionamiento de muchos aparatos electrónicos y mecánicos.

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

\section{Introduction}

Automatically recognizing facial expressions is an interesting and challenging problem in the fields human computer interaction and artificial emotional intelligence. Facial expressions can be interpreted and categorized into six classes of basic emotion that seem to be found across all human cultures: anger, disgust, fear, happiness, sadness, and surprise \cite{ekman93}. This projects goal is to develop a classifier for real-time detection of these facial expressions.

%------------------------------------------------

\section{Related work}
In the last years, great progress was made on the problem of automatically detecting facial expression from images. Some researchers have attempted to derive emotions from detecting individual muscle movements in images of faces by identifying muscles and their degree of activation \cite{lien98}. Manually extracting the relevant features from the training data requires expert knowledge or complex algorithms. In Convolutional Neural Networks (CNNs), features can be learned from the training data implicitly and only the image pixels are required as input for predictions. CNNs and traditional Support Vector Machines have been used for various image recognition tasks, including facial expression recognition \cite{dumas01}.
%------------------------------------------------

\section{Datasets}
To train an automatic classifier using deep learning technology, a dataset that approximates the problem is required. Each image in the dataset is mapped to an integer class label for the corresponding emotion. Two datasets, FER2013 \cite{goodfel13} and CK+ \cite{cohn00} are considered for training the deep learning model.

\begin{figure}[H]
	\includegraphics[width=0.48\textwidth]{fer_examples}
	\caption{FER-2013 training samples}
\end{figure}

\subsection{FER-2013}
The FER-2013 dataset consists of roughly 37.000 images showing faces of people expressing a certain emotion. Most of the 48x48 pixel grayscale images are stock footage. All images have been centered on the persons faces to show a similar region in every sample. As this has been done automatically, some images are not centered correctly. Persons are depicted from very different camera angles, some faces are occluded and rotation is different across the images. These problems make generalization of the training data difficult.

\begin{figure}[H]
	\includegraphics[width=0.48\textwidth]{ferplus_distribution}
	\caption{FER-2013 dataset class distribution}
\end{figure}

The class labels were crowd-sourced with Amazon Mechanical Turk, where quality work is not incentivized, so faulty labels are likely a problem. It is expected that every subject shows only a single emotion, where in reality, faces can express at least two emtions (e.g happiness and surprise) simultaniously. To resolve this problem and improve the overall quality of the FER-2013 dataset, a research group at Microsoft worked on re-labeling the dataset \cite{barsoum16}: To get a better approximation of the ground-truth emotions, ten workers on Mechanical Turk were asked to label each image. Instead of single class ground-truths, probability distributions of emotion classes describe each persons facial expression.

\subsection{Cohn-Kanade (CK+)}
To augment the wild variety of FER-2013 images, a second, more conservative dataset with only frontal faces is used to train a second classifier. The Cohn-Kanade (CK+) dataset consists of 593 image sequences across 123 subjects. Each image sequence goes from a neutral facial expression to a posed "peak expression" according to the six basic emotions. Only 327 of the 593 image sequences are labeled with emotion data. This is because not all images fit the prototypic definition of the Facial Action Coding System (FACS). CK+ also comes with FACS data, where facial action units (e.g raised eyebrow) are identified along with their degree of activation. To show that extracted features are not strictly required for training an automatic classifier, the FACS data is not used as input.

\begin{figure}[H]
	\includegraphics[width=0.48\textwidth]{ckplus_examples}
	\caption{Posed facial expressions from the CK+ dataset}
\end{figure}

%------------------------------------------------

\section{Convolutional Neural Networks (CNNs)}
Convolutional Neural Networks are are a special kind of multi-layer artificial neural networks. CNNs have been around since the 1990s \cite{lecun98}, however they have lead to breakthrough results in practical small-scale application with computational power becoming more affordable in the last years. Convolutional Neural Networks can recognize visual patterns directly from pixel images with very little preprocessing.

Like a regular Artificial Neural Network, a CNN can be thought of as a function $f(x) = y$ where the input $x$ is an image of a face and $y$ is a probability distribution of the classes this image is predicted to belong to (e.g happiness, surprise, sadness). Input images are flattened in dimension from 48x48 pixels to a single 2304-dimensional vector. For training and predictions, multiple images can be fed into the neural network at once.

The model function is implemented as a multi-layer computational graph where transformations are applied to the input on every layer.

\subsection{CNN Components}

\subsubsection{Convolutional Layer}
The convolutional layer is the core building block of a CNN. Each layer is initialized with a set of $n$ random two-dimensional filter matrices (eg 3x3 pixels) and a pixel-stride. When an image is put into the convolutional layer, each of the filters is moved across the height and width of the image, calculating a dot product between the filter matrix and the current section of the image. The resulting matrices of dot products are called feature maps. To introduce non-linearity to the operation, a ReLu activation function $f(n) = max(0, n)$ is used on the feature-maps element-wise. The filters in the convolutional layers are trained using a backpropagation algorithm. Intuitively, filters activate when they see different types of visual features, such as vertical or diagonal edges on the lower layers or more complex patterns like eyes or mouths in the top convolutional layers.

\subsubsection{Pooling Layer}
To progressively reduce dimensionality and computational cost, pooling layers can be found in the computational graph. Similar to the convolutional layer, filters are moved across the layer input, executing a pooling operation on each position. A common pooling operation is taking the max value on every filter position (Max-Pooling). For instance when pooling with a filter size of 3x3 pixels and a stride of 2 is applied to a 48x48 pixels image, the pooling output is a 24x24 pixels image.

\subsubsection{Fully-Connected layers}
In CNNs, fully connected layers are added at the top of the network. The advanced features that have been extracted in the top convolutional layers are usually connected to two or more layers of weighted nodes. On the first fully-connected layer, linear combinations of the feature maps activations are created. On top of that, linear transformations are applied to reduceach layer transforms the output of the previouse the dimension to match the number of classes. A softmax function can be applied on the output layer to normalize the topmost fully connected layer into a valid probability distribution.

%------------------------------------------------

\section{Model Architecture and Training}
To effectively reduce dimensions between the input and output space and find a valid representation of the input image, a CNN is implemented using the TensorFlow higher-level APIs. The TensorFlow extimator API provides a convenient abstraction layer by encapsulating trainable classifiers.

In simple CNNs, it is common to combine convolutional and pooling layers to compute feature activations while reducing dimensionality. During the designing process, it turned out that stacking multiple convolutional layers in a VGG-Fashion \cite{simonyan14} improved the model accuracy on the test set by about 20 percent compared to alternating pooling and convolutional layers.

The model architecture has been improved incrementally throughout the designing process. Hyperparameter optimization has not been done programmatically but rather by trial-and-error.

%\subsection{AlexNet-like CNN}
%For the Imagenet Challenge (REF) Alex … proposed a deep neural network (DNN) model consisting of alternating convolution and pooling layers with local response normalization in between, followed by three fully connected layers.

\subsection{Training Procedure}
The CNN is trained on a subset of the FER-dataset according to the original train/test/validation split with the FERplus prabability distributions as labels. The models filters and weights/biases of the fully connected layers are fit to the training set using a backpropagation algorithm with the Adam-optimizer minimizing a cross-entropy loss function. To prevent overfitting the model on the training set, the loss value on the validation set (unseen data) is computed periodically during training. When the validation loss stops decreasing, training is stopped and the final model accuracy is computed on the test subset.

%------------------------------------------------

\section{Discussion and Conclusion}

% - CNN robustness to occlusion, distortions and geometric transformations.
The final model scored an accuracy of ~75\% on a test-set of 3557 samples. In live predictions with a webcam, the classifier has shown to be surprisingly accurate and is a cool gadget to play around with. However, predictions for the classes $disgust$ and $fear$ are hard to reproduce. This is most likely because the seven classes of emotion in the dataset are highly imbalanced. For instance disgust only has 500 samples while happy has much more. This results in biased predictions with higher probabilities of the majority classes. In order to balance the training dataset, data augmentation could be used: To increase the number of training examples for a given class, the images could be flipped horizontally or rotated in multiple random angles. For normalizing the images, it is a common practice to subtract the mean pixel intensities over the whole dataset from each picture. Given more time for this project, I would like to have experimented with such techniques of data augmentation.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

%\begin{thebibliography}{9}
%
%	% OpenCV Haarcascade Classifier
%	\bibitem{violajones}
%  	P. A. Viola and M. J. Jones. Robust real-time face detection.
%	International Journal of Computer Vision
%	57(2):137–154, 2004.
%
%	% the ICML 2013 workshop’s facial expression recognition challenge
%	\bibitem{replearning}
%	I. J. Goodfellow et al.
%	Challenges in representation learning: A report on three machine
%	learning contests.
%	Neural Networks
%	, 64:59 – 63, 2015.
%	Special Issue on ‘Deep Learning of Representations’
%	
%	%[LeCun and Bengio, 1995a]
%    %Y. LeCun and Y. Bengio. Convolutional networks for images, speech, and time-series. 
%    %In M. A. Arbib, 		editor, The Handbook of Brain Theory and Neural Networks. MIT Press, 1995
%    
%    % Paul Ekmans 1969 research paper "Facial Expressions"
%
% 
%\end{thebibliography}

\bibliography{paper}
\bibliographystyle{plain}

%----------------------------------------------------------------------------------------

\end{multicols}

\end{document}